\begin{englishabstract}

	Humans in the 21st century have been living in the Information Age with the Internet as the core. With its outstanding characteristics and advantages of ubiquity, convenience, and intelligence, mobile networks have become a new space for people to study, work, and live, as well as an emerging and dynamic field of economic and social development. At the same time, with mobile networks' prominent instrumental attribute, social attributes, media attributes, and platform attributes, people demand for high-quality mobile services, and have increasing concerns about privacy protection and security risks of mobile services. Therefore, as we rely on the full power of commercial 5G technique to improve the performance of mobile networks, it is increasingly important to study the mechanism design of innovative mobile networks and optimize its overall utility, taking into account critical constraints, including 1) the resources limitation in the operation of mobile networks, 2) the self-interest characteristic of individual network participants, 3) the conflict between enhancing network performance and data privacy protection, and 4) the prominent effects of network externality. 
	
	Researchers have made significant progress in studying the utility optimization and mechanism design in mobile networks, though there are still some deficiencies. For example, there is a lack of in-depth analysis and discussion from the perspective of network externalities in solving the problem of utility optimization; the consideration and characterization of the trade-off between optimizing network utility and individuals' privacy protection is insufficient; and existing incentive mechanisms are designed based on ideal rationality assumptions for the individuals' behaviors, which has limitations in its practical usage.
	
	Based on the existing works on related topics, this thesis focuses on improving the utilities of different entities in the mobile networks from the perspective of the network externality. In particular, utility optimization problems in three basic mobile networks scenarios, namely: the crowdsensing network, the spectrum sharing network, and the data services market, are studied by using theoretic tools of game theory and mechanism design, to exploit the individuals' intelligence and interactions. Accordingly, a privacy-preserving incentive mechanism, a privacy-preserving spectrum access mechanism and a data pricing mechanism are proposed as solutions to the above three problems. The results enrich the theory of mobile network utility optimization and shed light on the design and practice of mobile network. The main contributions of the thesis are summarized as follows: 

	%With the popularity of mobile devices and the rapid development of related fields such as mobile social networks and online entertainment, subscribed mobile users and data traffic around the world have experienced a trend of substantial growth in recent years. The enlarging demands for high-quality communication and data service give rises to new challenges to the development of wireless network technology. To optimize the overall network performance, the network operators can rely on the enhancement on network infrastructure and capabilities brought by the next-generation 5G technology. In addition, they ought to exploit the limited resource in an efficient and reasonable manner via sophisticated designed resource allocation mechanism and service pricing mechanism. On the other hand, with the growing sensing capabilities of mobile devices, the mobile network has become the main channel for third-party applications to collect users' personal data and urban environment data, which has promoted the development of mobile big data technologies. While on the other side of the coin, the exposure of personal data has incurred serious privacy crisis and raised people's concern about the risk of privacy leakage.
	
%	As one of the main topics in the research field of network economics, network externality is defined as the indirect impact on an individual's utility from other individuals in the rest of the network. Such kind of effect may simply depend on the size of the network, or depend on the degree of involvement of other individuals, as well as the strength of the association between individuals. In recent years, the continuous expansion on the network-scale has made the phenomenon of network externalities increasingly prominent in some scenarios. Specifically, the utility obtained by an individual user from the network can be increased (with positive network externality) or decreased (with negative network externality) by the enhanced involvement of other users in the network. It is due to such prominent influences that network externality has on the individuals' utility, taking the network externality into consideration is of great importance. 
%In recent years, researchers have made a lot of progress in the protocol designs aiming for network performance optimization and privacy protection. However, there are still three aspects that require more efforts to be taken: 1) the lack of analysis and investigation on solving network optimization from the view of network externality; 2) the lack of research work combining network performance optimization and users' privacy protection, especially the characterization on the trade-off between the system performance and the privacy; 3) the lack of heterogenous modeling about individuals' strategic behaviors. Based on the state-of-the-art, this thesis has made an effort on network performance optimization within three scenarios: the spectrum sharing network, the mobile crowdsensing network, and the mobile data service market. The main contributions are summarized as follows: 
	\begin{enumerate}
		%\item A brief introduction on the background of network economics and the basic concepts of network externality and some discussion on the related works about mechanism design and privacy protection in mobile networks, is provided.
    		%\item The first part studies the cost minimization problem of privacy-preserving mobile crowdsensing system and proposes an auction based framework for privacy-preserving data aggregation. Given the heterogeneity on mobile users' sensing capabilities and their unit privacy cost, the platform is facing the challenges of determining the set of mobile users for which the noise distribution need to be carefully designed so that the privacy protection is provided and the data aggregation accuracy is satisfied. This work carries out a local data perturbation scheme and reveal the negative externality concerning the users' privacy-preserving level. And two different settings, ``privacy passive'' setting and ``privacy proactive'' setting, are discussed separately which corresponds to two kinds of privacy-preserving attitudes. Based on the hidden monotonicity of the problem, two computational efficient incentive mechanisms, DPDA and EDPDA, are proposed satisfying the truthfulness and individual rationality properties. By using the the proposed algorithms, the platform can approximately minimize the cost under the aggregation accuracy constraint in both two settings. The performance of the algorithm is validated through extensive numerical experiments.
		
		\item Study the cost minimization problem of privacy-preserving mobile crowdsensing system under the single network externality effect. An auction based framework for privacy-preserving data aggregation is proposed while considering the conflict objectives of collecting data from mobile users and protect users' data privacy. Given the heterogeneity on mobile users' sensing capabilities and their unit privacy cost, this work carries out a local data perturbation scheme so that the privacy protection is provided and the data aggregation accuracy is satisfied. The \textsl{negative network externality} concerning the users' privacy-preserving level is then explicitly revealed. And two different settings, ``privacy passive'' setting and ``privacy proactive'' setting, are discussed separately which corresponds to two kinds of privacy-preserving attitudes. Based on the hidden monotonicity of the problem, two computational efficient incentive mechanisms, DPDA and EDPDA, are proposed satisfying the truthfulness and individual rationality properties. By using the the proposed algorithms, the platform can approximately minimize the cost under the aggregation accuracy constraint in both two settings. The performance of the algorithm is validated through extensive numerical experiments.
		
    		%\item The second part studies the socially-aware throughput maximization problem within the context of locational privacy-preserving database-assisted spectrum sharing. The mobile users jointly take into account their physical coupling (negative externality due to the signal interference) and social coupling (positive externality due to the social network effect) while making spectrum sharing decisions. In particular, to mitigate RSS-based PHY-layer location privacy threat, a power perturbation approach is employed where each secondary user judiciously ``reduces'' its transmission power by choosing a power level following a statistical distribution (with a negative bias). Accordingly, the privacy-preserving spectrum sharing among users is cast as a stochastic channel selection game, where strategic players (secondary users) adjust their strategies dynamically aiming to maximize their social group utilities. Specifically, a two-time-scale distributed learning algorithm based on no regret-based rule is devised, which is shown to converge almost surely towards the set of socially-aware correlated equilibrium. The numerical results corroborate that the higher the privacy protection level, the more significant the degradation of the network throughput would be.
		
		\item Study the throughput maximization problem under the dual network externality effects within the context of locational privacy-preserving database-assisted spectrum sharing and propose a distributed spectrum sharing mechanism. Firstly, to mitigate RSS-based PHY-layer location privacy threat, a power perturbation approach is employed where each secondary user judiciously ``reduces'' its transmission power by choosing a power level following a statistical distribution (with a negative bias). Then the double impacts of network externality are explicitly revealed: the mobile users jointly take into account their physical coupling (\textsl{negative network externality} due to the signal interference) and social coupling (\textsl{positive network externality} due to the social network effect) while making spectrum sharing decisions. Accordingly, the privacy-preserving spectrum sharing among users is cast as a stochastic channel selection game, where strategic players (secondary users) adjust their strategies dynamically aiming to solve a social group utility maximization (SGUM) problem. Specifically, a two-time-scale distributed learning algorithm based on no regret-based rule is devised, which is shown to converge almost surely towards the set of socially-aware correlated equilibrium. The numerical results corroborate that the higher the privacy protection level, the more significant the degradation of the network throughput would be.
    		
		%\item The third part studies the revenue maximization problem for the wireless service providers. The pricing strategies of multiple service providers in a competitive data service market are studied where mobile users' data consumption behaviors are influenced by two effects: the network effect (positive network externality) and the congestion effect (negative network externality). To analyze the strategic interactions between mobile users and service providers, a two-stage Stackelberg game is devised, consisting of a providers' game in Stage I and a users' game in Stage II, respectively. In particular, for the users' game, the equilibrium solution is characterized explicitly and its uniqueness is established. For the providers' game, the analysis indicates that a mixed-strategy equilibrium solution is guaranteed for the scenario with rational providers as well as the scenario with providers of bounded rationality. A distributed learning algorithm for finding a mixed-strategy equilibrium solution is further provided. And the numerical results provide insights into how positive network effect and congestion effect would impact the system performance, and demonstrate that the bounded rational behavior incurs degradation to service providers' revenues.
		
		\item Study the revenue maximization problem for the wireless service providers in a competitive data service market under the dual network externality effects and propose a pricing mechanism based on Stackelberg game model. 
Firstly, the two effects that influence the mobile users' data consumption behaviors are studied, namely: the network effect (\textsl{positive network externality}) and the congestion effect (\textsl{negative network externality}). To analyze the strategic interactions between mobile users and service providers, a two-stage Stackelberg game is devised, consisting of a providers' game in Stage I and a users' game in Stage II, respectively. In particular, for the users' game, the equilibrium solution is characterized explicitly and its uniqueness is established. For the providers' game, the analysis indicates that a mixed-strategy equilibrium solution is guaranteed for the scenario with rational providers as well as the scenario with providers of bounded rationality. A distributed learning algorithm for finding a mixed-strategy equilibrium solution is further provided. And the numerical results provide insights into how positive network effect and congestion effect would impact the system performance, and demonstrate that the bounded rational behavior incurs degradation to service providers' revenues.
    		%\item The conclusions are drawn with future work at the end of the dissertation.
		
		\item Explore the methodology of integrating network economics into the research of mobile networking technology, and innovatively study the utility optimization problem of mobile network from the perspective of network externality. This thesis chooses three independent mobile network scenarios, reveals the network externalities and proposes network utility optimization solutions accordingly. The results have shown that a deep understanding and accurate characterization of the network externality of mobile networks is necessary and indispensable for enhancing the feasibility of mechanism design and the effectiveness of utility optimization. Moreover, the results turn out to be an extension and enrichment of the research on the applications of network economic theory. In addition, the theoretical tools of game theory and mechanism design used in this thesis combine the cutting-edge results in related fields with mature theoretical models, and also pay attention to their applicability in the real scenarios of mobile networks.
	\end{enumerate}

	\englishkeywords{Mobile Network, Utility Optimization, Mechanism Design, Network Externality, Game Theory}

\end{englishabstract}
